---
title: "Lab 11 - Smoking during pregnancy"
author: "Lindsey Wilson"
date: "5/3/23"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse) 
library(tidymodels)
library(openintro)
```

```{r load-data-set-seed}
nc <- ncbirths
set.seed(85920305)
```
## Part 1

### Exercise 1

There are 1000 cases in the dataset, where each case represents one baby born in North Carolina in 2004

### Exercise 2

Let's filter our dataset to only include babies born to White mothers and calculate their average birth weight:

```{r filter-white}
ncbirths_white <- nc %>%
  filter(whitemom == "white")

mean(ncbirths_white$weight)
```

Looks like the average White baby (or at least, the average baby born to a White mother) weighed 7.25. pounds at birth in 2004

### Exercise 3

I think so.  I'm not totally sure what this is asking,  but as far as I can tell the observations are independent (for example, nothing would allow us to tell whether individuals in the dataset are related in any way)

### Exercise 4

Let's actually run our test using bootstrapping:

```{r bootstrap}
df_boot_weight <- ncbirths_white %>%
  specify(response = weight) %>%
  generate(reps = 15000, type = "bootstrap") %>%
  calculate(stat = "mean")

df_boot_weight <- df_boot_weight %>%
  mutate(stat = stat + (7.43 - mean(ncbirths_white$weight)))

df_boot_weight %>%
  ggplot(aes(x = stat)) + 
  geom_histogram(color = "black", fill = "blue")


upper_tail <- sum(df_boot_weight$stat >= 7.61)/length(df_boot_weight$stat)
lower_tail <- sum(df_boot_weight$stat <= 7.25)/length(df_boot_weight$stat)
p_val <- upper_tail + lower_tail
```
Based on this calculation, it looks like the 2-tailed p-value for our observed sample mean of 7.25 is .00053, which is much less than .05. Therefore, it looks like our observed sample of White babies in 2004 weighed significantly less than White babies in 1995.

## Part 2

### Exercise 5

Let's visualize weight for smokers vs. nonsmokers:

```{r weight-v-habit}
nc %>%
  drop_na(habit) %>%
  ggplot(aes(x= habit, y = weight)) +
  geom_boxplot(fill = "blue")
```

What we can see from these plots is that:
1.)  The mean birth weight of babies born to smokers and nonsmokers isn't very different,  and
2.)  There is a ton of variability in both groups

### Exercise 6

Before moving on, let's clean the dataset more permanently by creating a new dataframe where values of "NA" for `habit` and `weight` are filtered out:

```{r filter-NA}
ncbirths_clean <- nc %>%
  drop_na(habit) %>%
  drop_na(weight)
```

### Exercise 7

Now let's calculate the difference in weight between smokers and nonsmokers

```{r summarize}
ncbirths_clean %>%
  group_by(habit) %>%
  summarise(mean_weight = mean(weight))
```

This makes the difference in mean weight between the groups 0.316 pounds, where babies of nonsmokers weigh slightly more.

### Exercise 8

H0: Babies born to smokers do not differ in weight from babies born to nonsmokers

H1: Babies born to smokers differ in weight from babies born to nonsmokers

### Exercise 9

Now let's actually run the hypothesis test:

```{r bootstrap-weight-smoking}
#Smokers
df_boot_smoker <- ncbirths_clean %>%
  filter(habit == "smoker") %>%
  specify(response = weight) %>%
  generate(reps = 15000, type = "bootstrap") %>%
  calculate(stat = "mean")
  
#Nonsmokers
df_boot_nonsmoker <- ncbirths_clean %>%
  filter(habit == "nonsmoker") %>%
  specify(response = weight) %>%
  generate(reps = 15000, type = "bootstrap") %>%
  calculate(stat = "mean")
  
#Take the difference between groups
df_boot_diff <- (df_boot_nonsmoker$stat - df_boot_smoker$stat) %>%
  as.data.frame() %>%
  setNames("MeanDiff")

#Calculate deviations
df_boot_diff_devs <- (df_boot_diff$MeanDiff - mean(df_boot_diff$MeanDiff)) %>%
  as.data.frame() %>%
  setNames("MeanDev")

#Calculate p value
upper_tail_2 <- (sum(df_boot_diff_devs$MeanDev >= mean(df_boot_diff$MeanDiff))/length(df_boot_diff$MeanDiff))

lower_tail_2 <- (sum(df_boot_diff_devs$MeanDev <= mean(df_boot_diff$MeanDiff)*-1)/length(df_boot_diff$MeanDiff))

p_val_2 <- upper_tail_2 + lower_tail_2
p_val_2
```

And the p-value I got here was .0157. Therefore,  it does look like there's a significant difference in weight between babies born to moms who smoke versus those who don't.

### Exercise 10

And let's add a 95% confidence interval for the mean difference in birth weight between babies born to smokers vs. nonsmokers:

```{r confidence-interval}
df_boot_diff %>%
  summarise(lower = quantile(MeanDiff, 0.025),
            upper = quantile(MeanDiff, 0.975))
```
It looks like zero does not fall within the interval. This again suggests that there's a real difference in birth weight between babies of smokers vs. nonsmokers.

## Part 3

### Exercise 11

Let's see what the age cutoffs are for the categories in the `mature` variable

```{r age-cutoff}
nc %>%
  group_by(mature) %>%
  summarise(min(mage), max(mage))
```

So it. looks like "younger mom" category ranges from age 13 to 34, and "older mom" covers any mother 35 and over.

### Exercise 12

Now let's run a hypothesis test comparing the proportion of low birthweight mothers between younger and mature moms (I'd like to give Conor credit for helping me out on this; I used his code as a template after I got stuck on this)

Hypotheses:

-  H0: the proportion of low birthweight mothers does not differ between younger and mature moms.

-  H1: the proportion of low birthweight mothers does differ between younger and mature moms.

```{r birthweight-test}
#Generate data sets
df_lo_mature <- ncbirths %>%
  filter(mature == "mature mom")

df_lo_young <- ncbirths %>%
  filter(mature == "younger mom") 

#Observed values
lo_mature <- sum((df_lo_mature$lowbirthweight == "low")/(nrow(df_lo_mature)))
lo_young <- sum((df_lo_young$lowbirthweight == "low")/(nrow(df_lo_young)))
obs <- lo_mature - lo_young

#Bootstrap 
boot_mature <- df_lo_mature %>% 
  specify(response = lowbirthweight, success = "low") %>% 
  generate(reps = 15000, type = "bootstrap") %>% 
  calculate(stat = "prop")

boot_young <- df_lo_young %>% 
  specify(response = lowbirthweight, success = "low") %>% 
  generate(reps = 15000, type = "bootstrap") %>%
  calculate(stat = "prop")

prop_diff <- (boot_mature$stat - boot_young$stat) %>% 
  as.data.frame() %>% 
  setNames("stat")

#Run Test 
prop_diff_c <- (prop_diff$stat - mean(prop_diff$stat)) %>% 
  as.data.frame() %>%
  #re-center at null
  setNames("stat")
sum((prop_diff_c$stat >= obs)/(nrow(prop_diff_c)))
```

Therefore, it looks like we don't have evidence of difference in the proportion of low birthweight babies between young and mature moms. 

### Exercise 13

The 95% confidence interval for the difference in the proportion of low birthweight babies born to young vs. mature moms is calculated below:

```{r birthweight-confidence-interval}
prop_diff %>%
  summarise(lower = quantile(stat, 0.025),
            upper = quantile(stat, 0.975))
```

We can be 95% confident that the true difference in low birthweight proportion between young and mature moms falls between -0.031 and 0.091. Since zero falls within this range, we don't have evidence to suggest there's a difference between the two groups.


